---
title: "A_model"
author: "Yiying Chen yc4776"
date: "2026-03-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r model_A_prep}
load("clean_datA.RData")
train <- train_data
test <- test_data
```

```{r multiple_linear_regression_variable}
fit_lm <- lm(log_PFS ~ age + menopause + race + bmi + tumor_size + tumor_grade +
               lymph_nodes + ER + PR + HER2 + Ki67 + stage + surgery +
               chemo + hormonal_therapy + lymph_nodes_log, data = train)
summary(fit_lm)
```
# From the summary, we get "age", "tumor_size", "lymph_nodes", "ERPositive", "HER2Positive", "hormonal_therapyYes" as significant variables, and "tumor_grade.L" with "Ki67" as marginally significant variables.
```{r lm_model}
#lm with only significant variables
fit_lm_significant_1 <- lm(log_PFS ~ age + tumor_size + lymph_nodes + ER + HER2 + hormonal_therapy, data = train)
#lm with significant variables + tumor_grade
fit_lm_significant_2 <- lm(log_PFS ~ age + tumor_size + lymph_nodes + ER + HER2 + hormonal_therapy + tumor_grade, data = train)
#lm with significant variables + Ki67
fit_lm_significant_3 <- lm(log_PFS ~ age + tumor_size + lymph_nodes + ER + HER2 + hormonal_therapy + Ki67, data = train)
#lm with significant variables + tumor_grade + Ki67
fit_lm_significant_4 <- lm(log_PFS ~ age + tumor_size + lymph_nodes + ER + HER2 + hormonal_therapy + tumor_grade + Ki67, data = train)

#observe the summary result
summary(fit_lm_significant_1)
summary(fit_lm_significant_2)
summary(fit_lm_significant_3)
summary(fit_lm_significant_4)

#observe the RMSE to decide the final model
pred1 <- predict(fit_lm_significant_1, newdata = test)
pred2 <- predict(fit_lm_significant_2, newdata = test)
pred3 <- predict(fit_lm_significant_3, newdata = test)
pred4 <- predict(fit_lm_significant_4, newdata = test)


rmse1 <- sqrt(mean((test$log_PFS - pred1)^2))
rmse2 <- sqrt(mean((test$log_PFS - pred2)^2))
rmse3 <- sqrt(mean((test$log_PFS - pred3)^2))
rmse4 <- sqrt(mean((test$log_PFS - pred4)^2))


rmse1
rmse2
rmse3
rmse4
```
# Based on the test error result, we should choose the model with all significant variables and tumor grade as the best prediction model for log_PFS, with RMSE 0.6853503

```{r GAM}
library(mgcv)

fit_gam <- gam(log_PFS ~ s(age) + s(tumor_size) + s(lymph_nodes) + s(Ki67) +
                 ER + HER2 + hormonal_therapy + tumor_grade, data = train)

summary(fit_gam)

pred_gam <- predict(fit_gam, newdata = test)

rmse_gam <- sqrt(mean((test$log_PFS - pred_gam)^2))
rmse_gam
```
# GAM gave only negligible improvement in RMSE, and most smooth terms were approximately linear, with only mild evidence of nonlinearity for tumor size.

```{r MARS, warning = FALSE, message = FALSE}
library(earth)

fit_mars <- earth(log_PFS ~ age + tumor_size + lymph_nodes + ER + HER2 + hormonal_therapy
                  + tumor_grade + Ki67, data = train)

summary(fit_mars)
pred_mars <- predict(fit_mars, newdata = test)
rmse_mars <- sqrt(mean((test$log_PFS - pred_mars)^2))
rmse_mars
```
# Based on the RMSE and interpretability, lm is also a better model than MARS.

```{r ridge_regression}
library(glmnet)
set.seed(1)

x_train <- model.matrix(log_PFS ~ age + tumor_size + lymph_nodes + ER + HER2 +
                          hormonal_therapy + tumor_grade + Ki67, data = train)[, -1]
y_train <- train$log_PFS

x_test <- model.matrix(log_PFS ~ age + tumor_size + lymph_nodes + ER + HER2 + hormonal_therapy
                       + tumor_grade + Ki67, data = test)[, -1]
y_test <- test$log_PFS

cv.ridge <- cv.glmnet(x_train, y_train, alpha = 0)
cv.ridge$lambda.min

pred_ridge <- predict(cv.ridge, newx = x_test, s = "lambda.min")
rmse_ridge <- sqrt(mean((y_test - pred_ridge)^2))
rmse_ridge

coef(cv.ridge, s = "lambda.min")
```
# The ridge regression model has a RMSE of 0.6856145, which is a bit larger than multiple linear regression model.

```{r lasso}
set.seed(1)
cv.lasso <- cv.glmnet(x_train, y_train, alpha = 1)
cv.lasso$lambda.min

pred_lasso <- predict(cv.lasso, newx = x_test, s = "lambda.min")
rmse_lasso <- sqrt(mean((y_test - pred_lasso)^2))
rmse_lasso

coef(cv.lasso, s = "lambda.min")
```
# The lasso model has a RMSE of 0.6853668, which is slightly larger than multiple linear regression model.

```{r elastic_net, message = FALSE}
library(caret)

set.seed(1)

ctrl1 <- trainControl(method = "cv", number = 10)

enet.fit <- train(
  log_PFS ~ age + tumor_size + lymph_nodes + ER + HER2 + hormonal_therapy + tumor_grade + Ki67,
  data = train,
  method = "glmnet",
  tuneGrid = expand.grid(
    alpha = seq(0, 1, length = 21),
    lambda = exp(seq(6, -4, length = 100))
  ),
  trControl = ctrl1
)

enet.fit$bestTune

enet.pred <- predict(enet.fit, newdata = test)
enet_test_rmse <- sqrt(mean((enet.pred - test$log_PFS)^2))
enet_test_rmse

coef(enet.fit$finalModel, enet.fit$bestTune$lambda)
```
# The elastic net model has a RMSE of 0.6856545 and lamda = 0, which means it is actually same as ridge regression and it is a bit larger than multiple linear regression model.