---
title: "A_model"
author: "Yiying Chen yc4776"
date: "2026-03-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r model_A_prep}
load("clean_datA.RData")
train <- train_data
test <- test_data
full_formula <- log_PFS ~ age + menopause + race + bmi + tumor_size + tumor_grade + lymph_nodes_log + ER + PR + HER2 + Ki67 + stage + surgery + chemo + hormonal_therapy
```

```{r multiple_linear_regression_variable}
fit_lm <- lm(full_formula, data = train)
summary(fit_lm)
```
# From the summary, we get "age", "tumor_size", "lymph_nodes_log", "ERPositive", "HER2Positive", "hormonal_therapyYes" as significant variables, and "tumor_grade.L" with "Ki67" as marginally significant variables.
```{r lm_model}
# lm with only significant variables
fit_lm_significant_1 <- lm(log_PFS ~ age + tumor_size + lymph_nodes_log + ER + HER2 + hormonal_therapy, data = train)
# lm with significant variables + tumor_grade
fit_lm_significant_2 <- lm(log_PFS ~ age + tumor_size + lymph_nodes_log + ER + HER2 + hormonal_therapy + tumor_grade, data = train)
# lm with significant variables + Ki67
fit_lm_significant_3 <- lm(log_PFS ~ age + tumor_size + lymph_nodes_log + ER + HER2 + hormonal_therapy + Ki67, data = train)
# lm with significant variables + tumor_grade + Ki67
fit_lm_significant_4 <- lm(log_PFS ~ age + tumor_size + lymph_nodes_log + ER + HER2 + hormonal_therapy + tumor_grade + Ki67, data = train)

# observe the summary result
summary(fit_lm_significant_1)
summary(fit_lm_significant_2)
summary(fit_lm_significant_3)
summary(fit_lm_significant_4)

# observe the RMSE to decide the final model
pred1 <- predict(fit_lm_significant_1, newdata = test)
pred2 <- predict(fit_lm_significant_2, newdata = test)
pred3 <- predict(fit_lm_significant_3, newdata = test)
pred4 <- predict(fit_lm_significant_4, newdata = test)


rmse1 <- sqrt(mean((test$log_PFS - pred1)^2))
rmse2 <- sqrt(mean((test$log_PFS - pred2)^2))
rmse3 <- sqrt(mean((test$log_PFS - pred3)^2))
rmse4 <- sqrt(mean((test$log_PFS - pred4)^2))


rmse1
rmse2
rmse3
rmse4
```
# Based on the test error result, we should choose the model with all significant variables and tumor grade as the best prediction model for log_PFS, with RMSE 0.6853503

```{r GAM}
library(mgcv)

fit_gam <- gam(log_PFS ~ s(age) + s(tumor_size) + s(lymph_nodes_log) + s(Ki67) +
                 ER + HER2 + hormonal_therapy + tumor_grade, data = train)

summary(fit_gam)

pred_gam <- predict(fit_gam, newdata = test)

rmse_gam <- sqrt(mean((test$log_PFS - pred_gam)^2))
rmse_gam
```
# GAM gave only negligible improvement in RMSE, and most smooth terms were approximately linear, with only mild evidence of nonlinearity for tumor size.

```{r MARS, warning = FALSE, message = FALSE}
library(earth)

fit_mars <- earth(log_PFS ~ age + tumor_size + lymph_nodes_log + ER + HER2 + hormonal_therapy
                  + tumor_grade + Ki67, data = train)

summary(fit_mars)
pred_mars <- predict(fit_mars, newdata = test)
rmse_mars <- sqrt(mean((test$log_PFS - pred_mars)^2))
rmse_mars
```
# Based on the RMSE and interpretability, lm is also a better model than MARS.

```{r linear_model_prep}
library(glmnet)

x_train <- model.matrix(full_formula, data = train)[, -1]
y_train <- train$log_PFS

x_test <- model.matrix(full_formula, data = test)[, -1]
y_test <- test$log_PFS
```

```{r ridge_regression}
set.seed(1)
cv.ridge <- cv.glmnet(x_train, y_train, alpha = 0)

cv.ridge$lambda.min
cv.ridge$lambda.1se

# test RMSE using best
pred_ridge_best <- predict(cv.ridge, newx = x_test, s = "lambda.min")
rmse_ridge_best <- sqrt(mean((y_test - pred_ridge_best)^2))
rmse_ridge_best

# test RMSE using 1se
pred_ridge_1se <- predict(cv.ridge, newx = x_test, s = "lambda.1se")
rmse_ridge_1se <- sqrt(mean((y_test - pred_ridge_1se)^2))
rmse_ridge_1se

# get the coefficients to see which ones are non-zero
coef(cv.ridge, s = "lambda.min")
coef(cv.ridge, s = "lambda.1se")
```
# For ridge regression, we use the full formula and get the RMSE of 0.6886818 from best fitted model.

```{r lasso}
set.seed(1)
cv.lasso <- cv.glmnet(x_train, y_train, alpha = 1)

cv.lasso$lambda.min
cv.lasso$lambda.1se

# test RMSE using best
pred_lasso_best <- predict(cv.lasso, newx = x_test, s = "lambda.min")
rmse_lasso_best <- sqrt(mean((y_test - pred_lasso_best)^2))
rmse_lasso_best

# test RMSE using 1se
pred_lasso_1se <- predict(cv.lasso, newx = x_test, s = "lambda.1se")
rmse_lasso_1se <- sqrt(mean((y_test - pred_lasso_1se)^2))
rmse_lasso_1se

# get the coefficients to see which ones are non-zero
coef(cv.lasso, s = "lambda.min")
coef(cv.lasso, s = "lambda.1se")
```
# For lasso regression, the best model had a test RMSE of 0.6869, while the 1SE model  had a higher RMSE of 0.6977. The 1SE model kept fewer predictors and was more parsimonious, but its predictive performance was worse. Therefore, the best lasso model performed better than the 1SE lasso model, although neither outperformed the multiple linear regression model.

```{r elastic_net, message = FALSE}
library(caret)

set.seed(1)

ctrl1 <- trainControl(method = "cv", number = 10)

enet.fit <- train(full_formula, data = train, method = "glmnet", 
                  tuneGrid = expand.grid(alpha = seq(0, 1, length = 21),
                  lambda = 10^seq(-4, 1, length = 100)),
                  trControl = ctrl1)
# selected tuning parameter
enet.fit$bestTune

# get alpha = 0.3
set.seed(1)
cv.enet <- cv.glmnet(x_train, y_train, alpha = 0.3)

cv.enet$lambda.min
cv.enet$lambda.1se

# test RMSE using best
pred_enet_best <- predict(cv.enet, newx = x_test, s = "lambda.min")
rmse_enet_best <- sqrt(mean((y_test - pred_enet_best)^2))
rmse_enet_best

# test RMSE using 1se
pred_enet_1se <- predict(cv.enet, newx = x_test, s = "lambda.1se")
rmse_enet_1se <- sqrt(mean((y_test - pred_enet_1se)^2))
rmse_enet_1se

# get the coefficients to see which ones are non-zero
coef(cv.enet, s = "lambda.min")
coef(cv.enet, s = "lambda.1se")
```
# In elastic net regression, the best model had lower test RMSE than the 1SE model. Although the 1SE model was simpler, it predicted less accurately. Neither elastic net model performed better than the multiple linear regression model.