---
title: "Group 22 Midterm Project"
date: "2026-03-25"
output:
  pdf_document:
    toc: true
    toc_depth: 3
  html_document:
    toc: true
    toc_float: true
---

\newpage

```{r setup, message = FALSE, warning = FALSE}
library(tidyverse)
library(caret)
library(glmnet)
library(mgcv)
library(earth)
library(pls)
library(patchwork)
library(vip) 
library(dplyr)
library(tidyr)
```

In order to reproduce the code we set the seed at 1
```{r}
set.seed(1) 
load("clean_datA.RData")
train <- train_data
test <- test_data
```

we performed 10-fold cross-validation exclusively within the training set
```{r}
ctrl <- trainControl(method = "cv", number = 10, selectionFunction = "best")
```

Include the `lymph_nodes_log` variable we created 
```{r}
full_formula <- log_PFS ~ age + menopause + race + bmi + tumor_size + tumor_grade +
  lymph_nodes_log + ER + PR + HER2 + Ki67 + stage + surgery + chemo + hormonal_therapy
```

# 1. Linear Regression (Baseline)
```{r}
set.seed(1)
fit_lm <- train(full_formula, data = train_data, method = "lm", trControl = ctrl)
```

# 2. Ridge, Lasso, Elastic Net (Regularization)

## 2-1. Ridge Regression
```{r}
set.seed(1)
fit_ridge <- train(full_formula, data = train_data, method = "glmnet",
                  tuneGrid = expand.grid(alpha = 0,
                                         lambda = exp(seq(-5, 5, length = 50))),
                  trControl = ctrl)

plot(fit_ridge, xTrans = log)

fit_ridge$bestTune
```


## 2-2. Lasso Regression
```{r warning = FALSE}
set.seed(1)
fit_lasso <- train(full_formula, data = train_data, method = "glmnet",
                  tuneGrid = expand.grid(alpha = 1,
                                         lambda = exp(seq(-8, 0, length = 50))),
                  trControl = ctrl)

plot(fit_lasso, xTrans = log)

fit_lasso$bestTune

coef(fit_lasso$finalModel, fit_lasso$bestTune$lambda)
```

## 2-3. Elastic net Regression
```{r warning = FALSE}
set.seed(1)
fit_enet <- train(full_formula, data = train_data, method = "glmnet",
                  tuneGrid = expand.grid(alpha = seq(0, 1, length = 11),
                                         lambda = exp(seq(-5, 1, length = 50))),
                  trControl = ctrl)

fit_enet$bestTune

plot(fit_enet, xTrans = log)

coef(fit_enet$finalModel, fit_enet$bestTune$lambda)
```

# 3. PCR & PLS (Dimension Reduction)
## 3-1. PCR (Principal Component Regression)
```{r}
set.seed(1)
pcr_mod <- pcr(full_formula, data = train_data, scale = TRUE, validation = "CV")

summary(pcr_mod)


fit_pcr <- train(full_formula, data = train_data, method = "pcr", 
                 scale = TRUE, trControl = ctrl, tuneLength = 15)
```


## 3-2. PLS (Partial Least Squares)
```{r}
set.seed(1)
fit_pls <- train(full_formula, data = train_data, method = "pls", 
                 scale = TRUE, trControl = ctrl, tuneLength = 15)
```

# 4. GAM & MARS (Nonlinear Methods)

## 4-1. GAM

```{r}
set.seed(1)
fit_gam <- train(full_formula, data = train_data, method = "gam", trControl = ctrl)

summary(fit_gam)

fit_gam$finalModel
```

## 4-2. MARS
```{r}
set.seed(1)
fit_mars <- train(full_formula, data = train_data, method = "earth", trControl = ctrl)

plot(fit_mars)
```

# Model comparison CV RMSE
```{r}
res_comparison <- resamples(list(LM = fit_lm, Ridge = fit_ridge, Lasso = fit_lasso, 
                                 ENet = fit_enet, PLS = fit_pls, PCR = fit_pcr, 
                                 GAM = fit_gam, MARS = fit_mars))
summary(res_comparison)

bwplot(res_comparison, metric = "RMSE", main = "CV RMSE Comparison")
```


# Test RMSE 
```{r}
model_list <- list(LM = fit_lm, Ridge = fit_ridge, Lasso = fit_lasso, 
                   ENet = fit_enet, PLS = fit_pls, PCR = fit_pcr, 
                   GAM = fit_gam, MARS = fit_mars)


test_preds <- as.data.frame(lapply(model_list, predict, newdata = test_data))


test_preds$log_PFS <- test_data$log_PFS


test_results <- test_preds |> 
  pivot_longer(cols = -log_PFS, names_to = "Model", values_to = "Predicted")  |> 
  group_by(Model)  |> 
  summarize(
    Test_RMSE = sqrt(mean((log_PFS - Predicted)^2)),
  )  |> 
  arrange(Test_RMSE) 

print(test_results)
```

```{r}
# Variable Importance for GAM model
gam_imp <- varImp(fit_gam)

plot(gam_imp, top = 10, main = "Variable Importance: Final GAM Model")
```


```{r}
vip(fit_mars$finalModel, type = "nsubsets")
```

