---
title: "TRY"
author: "Rosie Kwon"
date: "2026-03-25"
output: html_document
---

```{r}
library(tidyverse)
library(caret)
library(glmnet)
library(mgcv)
library(earth)
library(pls)
library(patchwork)
library(vip) 
```

In order to reproduce the code we set the seed at 1
```{r}
set.seed(1) 
load("clean_datA.RData")
train <- train_data
test <- test_data
```

we performed 10-fold cross-validation exclusively within the training set
```{r}
ctrl <- trainControl(method = "cv", number = 10, selectionFunction = "best")
```

Include the `lymph_nodes_log` variable we created 
```{r}
full_formula <- log_PFS ~ age + menopause + race + bmi + tumor_size + tumor_grade + lymph_nodes_log + ER + PR + HER2 + Ki67 + stage + surgery + chemo + hormonal_therapy
```

# (1) Linear Regression (Baseline)
```{r}
set.seed(1)
fit_lm <- train(full_formula, data = train_data, method = "lm", trControl = ctrl)
```

# (2) Ridge, Lasso, Elastic Net (Regularization)
## (2-1) Ridge
```{r}
set.seed(1)
fit_ridge <- train(full_formula, data = train_data, method = "glmnet",
                  tuneGrid = expand.grid(alpha = 0,
                                         lambda = exp(seq(-5, 5, length = 50))),
                  trControl = ctrl)

plot(fit_ridge, xTrans = log)

fit_ridge$bestTune
```


## (2-2) Lasso
```{r}
set.seed(1)
fit_lasso <- train(full_formula, data = train_data, method = "glmnet",
                  tuneGrid = expand.grid(alpha = 1,
                                         lambda = exp(seq(-8, 0, length = 50))),
                  trControl = ctrl)

plot(fit_lasso, xTrans = log)

fit_lasso$bestTune

coef(fit_lasso$finalModel, fit_lasso$bestTune$lambda)
```

## (2-3) Elastic net
```{r}
set.seed(1)
fit_enet <- train(full_formula, data = train_data, method = "glmnet",
                  tuneGrid = expand.grid(alpha = seq(0, 1, length = 11),
                                         lambda = exp(seq(-5, 1, length = 50))),
                  trControl = ctrl)

fit_enet$bestTune

plot(fit_enet, xTrans = log)

coef(fit_enet$finalModel, fit_enet$bestTune$lambda)
```

# (3) PCR & PLS (Dimension Reduction)
## (3-1) PCR (Principal Component Regression)
```{r}
set.seed(1)
fit_pcr <- train(full_formula, data = train_data, method = "pcr", 
                 scale = TRUE, trControl = ctrl, tuneLength = 15)
```


## (3-2) PLS (Partial Least Squares)
```{r}
set.seed(1)
fit_pls <- train(full_formula, data = train_data, method = "pls", 
                 scale = TRUE, trControl = ctrl, tuneLength = 15)
```

# (4) GAM & MARS (Nonlinear Methods)
## (4-1) GAM
```{r}
set.seed(1)
fit_gam <- train(full_formula, data = train_data, method = "gam", trControl = ctrl)
```

## (4-2) MARS
```{r}
set.seed(1)
fit_mars <- train(full_formula, data = train_data, method = "earth", trControl = ctrl)
```

# Model comparison
```{r}
res_comparison <- resamples(list(LM = fit_lm, Ridge = fit_ridge, Lasso = fit_lasso, ENet = fit_enet, PLS = fit_pls, PCR = fit_pcr, GAM = fit_gam, MARS = fit_mars))
summary(res_comparison)

bwplot(res_comparison, metric = "RMSE", main = "CV RMSE Comparison")
```


