---
title: "Group 22 Midterm Project"
output:
  pdf_document:
    toc: true
    toc_depth: 3
  html_document:
    toc: true
    toc_float: true
---

\newpage

# 1. Data preprocessing

## Load packages and data

```{r setup, message = FALSE, warning = FALSE}
library(tidyverse)
library(GGally)
library(corrplot)
library(caret)
library(glmnet)
library(mgcv)
library(earth)
library(pls)
library(patchwork)
library(vip) 
```

```{r}
load("datA.RData")
dfA <- datA
```

## Basic Structure
```{r}
# Dimensions
dim(dfA)

# Summary
summary(dfA)
```

## Convert variables to factors
```{r}
# Convert nominal and ordinal variables to factors
dfA <- dfA |>
  mutate(
    # Nominal variables
    menopause = factor(menopause, levels = c(0,1), labels = c("Pre","Post")),
    race = factor(race, levels = c(1,2,3,4),
                  labels = c("White","Black","Asian","Hispanic")),
    surgery = factor(surgery, levels = c(0,1),
                     labels = c("BCS","Mastectomy")),
    chemo = factor(chemo, levels = c(0,1),
                   labels = c("No","Yes")),
    hormonal_therapy = factor(hormonal_therapy, levels = c(0,1),
                             labels = c("No","Yes")),
    ER = factor(ER, levels = c(0,1), labels = c("Negative","Positive")),
    PR = factor(PR, levels = c(0,1), labels = c("Negative","Positive")),
    HER2 = factor(HER2, levels = c(0,1), labels = c("Negative","Positive")),

    # Ordinal variables
    tumor_grade = factor(tumor_grade, levels = c(1,2,3),
                         labels = c("Well","Moderate","Poor"),
                         ordered = TRUE),

    stage = factor(stage, levels = c(1,2,3),
                   labels = c("Stage I","Stage II","Stage III"),
                   ordered = TRUE)
  )
```

## Missing data check
```{r}
# Count 
colSums(is.na(dfA))
```

# 2. EDA for Dataset A

## Distribution of continuous variables
```{r}
# Histograms
dfA |> 
  select(age, bmi, tumor_size, lymph_nodes, Ki67, log_PFS) |>
  pivot_longer(cols = everything()) |>
  ggplot(aes(x = value)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "black") +
  facet_wrap(~name, scales = "free") +
  theme_minimal()
```

## Frequency distribution of categorical variables
```{r}
plot_categorical <- function(data, vars, title_text) {
  plot_data <- data |>
    select(all_of(vars)) |>
    mutate(across(everything(), as.character)) |>
    pivot_longer(cols = everything()) |>
    group_by(name, value) |>
    summarise(count = n(), .groups = "drop") |>
    group_by(name) |>
    mutate(
      percent = count / sum(count),
      label = paste0(count, "\n(", round(percent * 100, 1), "%)")
    )

  ggplot(plot_data, aes(x = value, y = count)) +
    geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
    geom_text(
      aes(label = label),
      vjust = -0.3,       
      color = "black",    
      size = 3,
      lineheight = 0.9  
    ) +
    facet_wrap(~name, scales = "free", ncol = 3) +
    scale_y_continuous(expand = expansion(mult = c(0, 0.25))) + 
    coord_cartesian(clip = "off") + 
    theme_minimal() +
    labs(x = "Category", y = "Count", title = title_text) +
    theme(
      strip.text = element_text(face = "bold", size = 11, margin = margin(b = 10)),
      axis.text.x = element_text(angle = 45, hjust = 1),
      panel.grid.major.x = element_blank(),
      plot.margin = margin(t = 20, r = 10, b = 10, l = 10),
      legend.position = "none"
    )
}

vars1 <- c("stage", "tumor_grade", "ER", "PR", "HER2")
vars2 <- c("race", "menopause", "chemo", "hormonal_therapy", "surgery")

plot_categorical(dfA, vars1, "Distribution of Categorical Variables")
plot_categorical(dfA, vars2, "Distribution of Categorical Variables Cont.")
```


## Relationship with response variable
```{r}
# Categorical
dfA |>
  select(log_PFS, stage, tumor_grade, ER, PR, HER2,
         chemo, hormonal_therapy, surgery, menopause) |>
  mutate(across(-log_PFS, as.character)) |>
  pivot_longer(cols = -log_PFS) |>
  ggplot(aes(x = value, y = log_PFS)) +
  
  geom_boxplot(fill = "lightblue") +
  
  facet_wrap(~name, scales = "free_x", ncol = 5) +
  
  theme_minimal() +
  labs(
    x = "Category",
    y = "log_PFS",
    title = "Association Between Categorical Variables and log_PFS"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
    strip.text = element_text(size = 10)
  )

# Continuous
dfA |>
  select(log_PFS, age, bmi, tumor_size, lymph_nodes, Ki67) |>
  pivot_longer(cols = -log_PFS) |>
  ggplot(aes(x = value, y = log_PFS)) +
  geom_point(alpha = 0.3, color = "lightblue") +
  geom_smooth(method = "loess", se = FALSE, color = "blue") +
  facet_wrap(~name, scales = "free_x") +
  theme_minimal() +
  labs(
    x = "Value",
    y = "log_PFS",
    title = "Association Between Continuous Variables and log_PFS"
  )
```

## Correlation Check (Continuous)
```{r}
library(corrplot)
cont_vars <- dfA |>
  select(age, bmi, tumor_size, lymph_nodes, Ki67, log_PFS)

cor_matrix <- cor(cont_vars, use = "complete.obs")

corrplot::corrplot(
  cor_matrix,
  method = "circle",         
  type = "full",
  tl.col = "red",            
  tl.srt = 45,               
  addCoef.col = "black",     
  number.cex = 0.7,
  mar = c(0,0,1,0)
)
```

## Handle Skewness
```{r}
# Log transform lymph_nodes to handle right-skewness
dfA <- dfA |>
  mutate(
    lymph_nodes_log = log(lymph_nodes + 1)
  )
dfA |>
  select(lymph_nodes, lymph_nodes_log) |>
  pivot_longer(cols = everything()) |>
  ggplot(aes(x = value)) +
  geom_histogram(aes(y = after_stat(density)),
                 binwidth = 0.2,
                 fill = "steelblue",
                 color = "black") +
  geom_density(color = "red", linewidth = 1) +
  facet_wrap(~name, scales = "free") +
  theme_minimal() +
  labs(title = "Before vs After Log Transformation")
```

## Train Test Split
```{r}
set.seed(1)
train_index <- createDataPartition(dfA$log_PFS, p = 0.7, list = FALSE)
train_data <- dfA[train_index, ]
test_data  <- dfA[-train_index, ]
```


We loaded the cleaned dataset from clean_datA.RData, which had already been split into a training set (train_data, n = 1400) and a testing set (test_data, n = 600). To ensure reproducibility and match with previous data cleaning process, we set the random seed to 1 before model fitting.

```{r}
set.seed(1) 
train <- train_data
test <- test_data
```

All model development and tuning were performed exclusively within the training set using 10-fold cross-validation.
```{r}
#cross validation setting 
ctrl <- trainControl(method = "cv", number = 10, selectionFunction = "best")
```

We first created a variable to include all predictors as well as the "lymph_nodes_log" we created.
```{r}
#full list of predictors using the transformed 'lymph_nodes_log' instead of the original 'lymph_nodes'
full_formula <- log_PFS ~ age + menopause + race + bmi + tumor_size + tumor_grade +
  lymph_nodes_log + ER + PR + HER2 + Ki67 + stage + surgery + chemo + hormonal_therapy
```

# 3. Model Building 
## 3-1. Linear Regression (Baseline)
The baseline linear regression model was fit using the full set of predictors. In the baseline model, older age, larger tumor size, greater lymph node burden, HER2 positivity, and lack of hormonal therapy were associated with shorter log-transformed PFS, whereas ER positivity was associated with longer log-transformed PFS.
```{r}
set.seed(1) # Ensure reproducible CV folds
fit_lm <- train(full_formula, data = train, method = "lm", trControl = ctrl)
fit_lm$finalModel
```

## 3-2. Ridge, Lasso, Elastic Net (Regularization)

### 3-2-1. Ridge Regression
The ridge regression was fit using the full set of predictors with alpha = 0 based on definition. The ridge regression model will choose the best lambda value by itself.
```{r}
set.seed(1)
fit_ridge <- train(full_formula, data = train, method = "glmnet",
                  tuneGrid = expand.grid(alpha = 0,
                                         lambda = exp(seq(-5, 5, length = 50))),
                  trControl = ctrl)

#plot the ridge regression curve
plot(fit_ridge, xTrans = log)

#see the chosen lambda for ridge regression model
fit_ridge$bestTune
```
From the ridge regression plot, we can see that the performance of CV RMSE was pretty well with about 0.71 as the lowest.


### 3-2-2. Lasso Regression
The lasso regression was fit using the full set of predictors with alpha = 1 based on definition. The lasso regression model will choose the best lamda value by itself.
```{r warning = FALSE}
set.seed(1)
fit_lasso <- train(full_formula, data = train, method = "glmnet",
                  tuneGrid = expand.grid(alpha = 1,
                                         lambda = exp(seq(-8, 0, length = 50))),
                  trControl = ctrl)

#plot the lasso curve
plot(fit_lasso, xTrans = log)

#see the chosen lambda for lasso model
fit_lasso$bestTune

#observe which coefficients are zero
coef(fit_lasso$finalModel, fit_lasso$bestTune$lambda)
```
Moreover, from the result curve of lasso regression model, we can observe that there is a obvious downwards trend when the tuning parameter approaches to -4. The cv rmse is slightly smaller than 0.71 at parameter = -4.

### 3-2-3. Elastic net Regression

The elastic net regression was fit using the full set of predictors without choosen alpha. The elastic net regression model will choose the best lamda value and alpha value by itself.
```{r warning = FALSE}
set.seed(1)
fit_enet <- train(full_formula, data = train, method = "glmnet",
                  tuneGrid = expand.grid(alpha = seq(0, 1, length = 11),
                                         lambda = exp(seq(-5, 1, length = 50))),
                  trControl = ctrl)

#see the chosen lambda for elastic net model
fit_enet$bestTune

#plot the elastic net curve
plot(fit_enet, xTrans = log)

#observe which coefficients are zero
coef(fit_enet$finalModel, fit_enet$bestTune$lambda)
```
The output and the plot shows that when alpha = 0.3 and lambda = 0.05402181, the lowest cv rmse value is about 0.71


## 3-3. PCR & PLS (Dimension Reduction)
### 3-3-1. PCR (Principal Component Regression)
The PCR model was fit using the full set of predictors with standardized variables. It reduced the original predictors to a smaller number of principal components and selected the number of components by cross-validation.
```{r}
set.seed(1)
pcr_mod <- pcr(full_formula, data = train, scale = TRUE, validation = "CV")

summary(pcr_mod)

# We chose 15 because the total number of predictors (including dummy variables) is around 19
fit_pcr <- train(full_formula, data = train_data, method = "pcr", 
                 scale = TRUE, trControl = ctrl, tuneLength = 15)
ggplot(fit_pcr, highlight = TRUE)

```

The PCR tuning plot suggests that predictive performance improved as the number of components increased up to about 8, where the CV RMSE reached its minimum, about 0.71. Beyond that point, adding more components did not improve prediction.

### 3-3-2. PLS (Partial Least Squares)
The PLS model was also fit using the full set of standardized predictors, with the number of components selected by cross-validation. PLS constructs components using information from both the predictors and the response.
```{r warning=FALSE}
set.seed(1)
fit_pls <- train(full_formula, data = train, method = "pls", 
                 scale = TRUE, trControl = ctrl, tuneLength = 15)
ggplot(fit_pls, highlight = TRUE)
```
From the plot, we can observe that PLS reached its best performance with only 1 component. This suggests that PLS was more efficient in extracting components that were directly related to the outcome.


## 3-4. GAM & MARS (Nonlinear Methods)

### 3-4-1. GAM
The GAM model was fit using the full set of predictors to allow for possible nonlinear effects.
```{r}
set.seed(1)
fit_gam <- train(full_formula, data = train, method = "gam", trControl = ctrl)

summary(fit_gam)

fit_gam$finalModel

plot(fit_gam)
```
From the plot, we can see that the cv rmse was slightly smaller for the GAM without feature selection, indicating that keeping all predictors produced better predictive performance than using feature selection.


### 3-4-2. MARS
The MARS model was fit as a flexible nonlinear regression model using the full set of predictors. It automatically searched for piecewise linear relationships and selected model complexity through cross-validation.
```{r}
set.seed(1)
fit_mars <- train(full_formula, data = train, method = "earth", trControl = ctrl)

plot(fit_mars)
```
From the result, it shows that although the minimum CV RMSE was achieved at the largest number of terms, the difference from the model with 12 terms was negligible. This suggests that increasing model complexity beyond 12 terms did not meaningfully improve prediction.

# 4. Model comparison using CV-RMSE
Based on the cross-validated RMSE results, the GAM achieved the smallest mean RMSE among all candidate models. This indicates that, during model development within the training set, the GAM had the strongest overall predictive performance. Since cross-validation provides an internal estimate of out-of-sample prediction error, this result suggests that the GAM was the most promising model among the candidate approaches we considered.
```{r}
# obtain all cv rmse results from the models
res_comparison <- resamples(list(LM = fit_lm, Ridge = fit_ridge, Lasso = fit_lasso, 
                                 ENet = fit_enet, PLS = fit_pls, PCR = fit_pcr, 
                                 GAM = fit_gam, MARS = fit_mars))
summary(res_comparison)

#visualize the comparison between cv rmse results
bwplot(res_comparison, metric = "RMSE", main = "CV RMSE Comparison")
```


# 5. Test RMSE 

We additionally reported the test RMSE to describe how the selected model performed on the independent test set. Among the test RMSE results, the GAM also had the lowest value, indicating that it maintained the strongest predictive performance on unseen data. Although we are not using test RMSE for model selection, this result was consistent with the cross-validation findings and provided additional evidence that the GAM generalized well to the independent test set.

```{r}
# get all test RMSE and put them into a data frame
model_list <- list(LM = fit_lm, Ridge = fit_ridge, Lasso = fit_lasso, 
                   ENet = fit_enet, PLS = fit_pls, PCR = fit_pcr, 
                   GAM = fit_gam, MARS = fit_mars)

test_preds <- as.data.frame(lapply(model_list, predict, newdata = test_data))

test_preds$log_PFS <- test_data$log_PFS

# arrange the test rmse results into a table from ascending order
test_results <- test_preds |> 
  rename(MARS = y) |> 
  pivot_longer(cols = -log_PFS, names_to = "Model", values_to = "Predicted")  |> 
  group_by(Model)  |> 
  summarize(
    Test_RMSE = sqrt(mean((log_PFS - Predicted)^2)),
  )  |> 
  arrange(Test_RMSE) 

knitr::kable(test_results)
```

## Variable Importance plot
Although the GAM output showed that many predictors had effects that were approximately linear (edf = 1.00), the model also suggested that some important predictors exhibited nonlinear patterns. To further justify the use of a nonlinear model, we examined the variable importance results from the GAM. We found that the two most influential predictors were both variables with nonlinear effects. This helps explain why the GAM achieved the best predictive performance.
```{r}
# Variable Importance for GAM model
gam_imp <- varImp(fit_gam)

plot(gam_imp, top = 10, main = "Variable Importance: Final GAM Model")
```

In conclusion, among all candidate models, the GAM showed the best predictive performance based on the lowest cross-validated RMSE. The test RMSE was reported as a supplementary measure on the independent test set, and it was consistent with the cross-validation results, with the GAM model also achieving the lowest value. Overall, for researcher A’s data,
log_PFS~menopause+race+tumor_grade+ER+PR+HER2
+stage+surgery+chemo+hormonal_therapy+
s(lymph_nodes_log)+s(tumor_size)+s(bmi)+s(Ki67)+s(age) 

from the GAM model is the best predictive model for PFS.

# 6. External validation - Researcher B

```{r}
#load 2000–2005 cohort data
load("datB.RData")
dfB <- datB
```

## Convert variables to factors
```{r}
dfB <- dfB |>
  mutate(
    # Nominal variables
    menopause = factor(menopause, levels = c(0,1), labels = c("Pre","Post")),
    race = factor(race, levels = c(1,2,3,4),
                  labels = c("White","Black","Asian","Hispanic")),
    surgery = factor(surgery, levels = c(0,1),
                     labels = c("BCS","Mastectomy")),
    chemo = factor(chemo, levels = c(0,1),
                   labels = c("No","Yes")),
    hormonal_therapy = factor(hormonal_therapy, levels = c(0,1),
                             labels = c("No","Yes")),
    ER = factor(ER, levels = c(0,1), labels = c("Negative","Positive")),
    PR = factor(PR, levels = c(0,1), labels = c("Negative","Positive")),
    HER2 = factor(HER2, levels = c(0,1), labels = c("Negative","Positive")),

    # Ordinal variables
    tumor_grade = factor(tumor_grade, levels = c(1,2,3),
                         labels = c("Well","Moderate","Poor"),
                         ordered = TRUE),

    stage = factor(stage, levels = c(1,2,3),
                   labels = c("Stage I","Stage II","Stage III"),
                   ordered = TRUE)
  )
```


```{r}
summary(dfB)
colSums(is.na(dfB))
```

## Distribution of continuous variables
```{r}
# Histograms
dfB |> 
  select(age, bmi, tumor_size, lymph_nodes, Ki67, log_PFS) |>
  pivot_longer(cols = everything()) |>
  ggplot(aes(x = value)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "black") +
  facet_wrap(~name, scales = "free") +
  theme_minimal()
```

## Frequency distribution of categorical variables
```{r}
vars1 <- c("stage", "tumor_grade", "ER", "PR", "HER2")
vars2 <- c("race", "menopause", "chemo", "hormonal_therapy", "surgery")

plot_categorical(dfB, vars1, "Distribution of Categorical Variables")
plot_categorical(dfB, vars2, "Distribution of Categorical Variables Cont.")
```

## Handle Skewness
```{r}
# Log transform lymph_nodes
dfB <- dfB |>
  mutate(
    lymph_nodes_log = log(lymph_nodes + 1)
  )
dfB |>
  select(lymph_nodes, lymph_nodes_log) |>
  pivot_longer(cols = everything()) |>
  ggplot(aes(x = value)) +
  geom_histogram(aes(y = after_stat(density)),
                 binwidth = 0.2,
                 fill = "steelblue",
                 color = "black") +
  geom_density(color = "red", linewidth = 1) +
  facet_wrap(~name, scales = "free") +
  theme_minimal() +
  labs(title = "Before vs After Log Transformation")
```

## Applying the GAM model to 2000-2005 cohort data

```{r}
pred_B <- predict(fit_gam, newdata = dfB)

#RMSE value 
rmse_B <- sqrt(mean((dfB$log_PFS - pred_B)^2))
print(rmse_B)
```

The GAM model that we built to predict for PFS does not generalize well to the 2000–2005 cohort. The RMSE with 2000-2005 cohort data increased to 0.8401 compared to the original 0.686. This discrepancy suggests a temporal shift in patient outcomes. 


# 7. Building New Model for 2000-2005 cohort data

To build a robust model for the 2000s, we also employed a 70/30 split and a consistent seed (1) to ensure the new findings are comparable to the previous study's methodology.
```{r}
#split dataset
set.seed(1) 
train_index <- createDataPartition(dfB$log_PFS, p = 0.7, list = FALSE)
train_dataB <- dfB[train_index, ]
test_dataB  <- dfB[-train_index, ]
```


## 7-1. Linear Regression (Baseline)
```{r}
set.seed(1)
fit_lmB <- train(full_formula, data = train_dataB, method = "lm", trControl = ctrl)

summary(fit_lmB)
```

## 7-2. Ridge, Lasso, Elastic Net (Regularization)

### 7-2-1. Ridge Regression
```{r}
set.seed(1)
fit_ridgeB <- train(full_formula, data = train_dataB, method = "glmnet",
                  tuneGrid = expand.grid(alpha = 0,
                                         lambda = exp(seq(-5, 5, length = 50))),
                  trControl = ctrl)

plot(fit_ridgeB, xTrans = log)

fit_ridgeB$bestTune
```


### 7-2-2. Lasso Regression
```{r}
set.seed(1)
fit_lassoB <- train(full_formula, data = train_dataB, method = "glmnet",
                  tuneGrid = expand.grid(alpha = 1,
                                         lambda = exp(seq(-8, 0, length = 50))),
                  trControl = ctrl)

plot(fit_lassoB, xTrans = log)

fit_lassoB$bestTune

coef(fit_lassoB$finalModel, fit_lassoB$bestTune$lambda)
```

### 7-2-3. Elastic net Regression
```{r}
set.seed(1)
fit_enetB <- train(full_formula, data = train_dataB, method = "glmnet",
                  tuneGrid = expand.grid(alpha = seq(0, 1, length = 11),
                                         lambda = exp(seq(-5, 1, length = 50))),
                  trControl = ctrl)

fit_enetB$bestTune

plot(fit_enetB, xTrans = log)

coef(fit_enetB$finalModel, fit_enetB$bestTune$lambda)
```

## 7-3. PCR & PLS (Dimension Reduction)
### 7-3-1. PCR (Principal Component Regression)
```{r}
set.seed(1)
#fit the pcr model using the function train() in caret
fit_pcrB <- train(full_formula, data = train_dataB, method = "pcr", 
                 scale = TRUE, trControl = ctrl,
                 tuneGrid = data.frame(ncomp = 1:19))

summary(fit_pcrB)




#determine the optimal number of components using plot 
ggplot(fit_pcrB, highlight = TRUE) + theme_bw()
```


### 7-3-2. PLS (Partial Least Squares)
```{r}
set.seed(1)
#fit the pls model using plsr()
pls_mod <- plsr(full_formula, data = train_dataB, scale = TRUE, validation = "CV")

summary(pls_mod)

#With the consistent using caret function to fit the model, we fit the model using train() function in caret
fit_plsB <- train(full_formula, data = train_dataB, method = "pls", 
                 scale = TRUE, trControl = ctrl, 
                 tuneGrid = data.frame(ncomp = 1:19))
summary(fit_plsB)


#determine the optimal number of components using plot 
ggplot(fit_plsB, highlight = TRUE) + theme_bw()
```

The optimal number of components is 2 

## 7-4. GAM & MARS (Nonlinear Methods)

### 7-4-1. GAM

```{r}
set.seed(1)
#fit the gam model using caret
fit_gamB <- train(full_formula, data = train_dataB, method = "gam", trControl = ctrl)

summary(fit_gamB)

fit_gam$finalModel

#While caret provides an efficient pipeline for model comparison, it limits the flexibility of the mgcv engine. We conducted a manual GAM analysis for Researcher B using the s() function for continuous variables
gam.B1 <- gam(log_PFS ~ s(age) + menopause + race + s(bmi) + s(tumor_size) + tumor_grade + s(lymph_nodes_log) + ER + PR + HER2 + s(Ki67) + stage + surgery + chemo + hormonal_therapy, data = train_dataB)

summary(gam.B1)

plot(gam.B1)
```

s(Ki67) has an edf value of 1.000, showing that Ki67 has a perfectly linear relationship with log_PFS.

### 7-4-2. MARS

For the MARS model, we defined a tuning grid with degree ranging from 1 to 3 and nprune from 2 to 19. The range for nprune was chosen to encompass the full set of predictors.
```{r}
set.seed(1)
mars_grid <- expand.grid(degree = 1:3, nprune = 2:19)

fit_marsB <- train(full_formula, data = train_dataB, method = "earth", tuneGrid = mars_grid, trControl = ctrl)

ggplot(fit_marsB)

fit_marsB$bestTune

coef(fit_marsB$finalModel)
```

# 8. Model comparison CV RMSE

Comparing the models based on resampling results.
```{r}
# List of the models
res_comparison <- resamples(list(LM = fit_lmB, Ridge = fit_ridgeB, Lasso = fit_lassoB, ENet = fit_enetB, PLS = fit_plsB, PCR = fit_pcrB, GAM = fit_gamB, MARS = fit_marsB))
summary(res_comparison)

bwplot(res_comparison, metric = "RMSE", main = "CV RMSE Comparison")
```

In order to check what is the reason behind the discrepancy, we created a variable importance plot using the GAM model that was trained with the 2000-2005 cohort.

```{r}
#Variable Importance
vip_B <- varImp(fit_gamB)
plot(vip_B, top = 10, main = "Variable Importance: 2000-2005 Cohort (Researcher B)")
```

In the 1990-1995 cohort, lymph_nodes_log was the predominant predictor, whereas in the 2000-2005 cohort, its importance markedly decreased, giving way to treatment-related factors like hormonal_therapy and biological markers like Ki67 and tumor_grade. 

# 9. Test RMSE 
```{r}
model_list <- list(LM = fit_lmB, Ridge = fit_ridgeB, Lasso = fit_lassoB, 
                   ENet = fit_enetB, PLS = fit_plsB, PCR = fit_pcrB, 
                   GAM = fit_gamB, MARS = fit_marsB)


test_preds <- as.data.frame(lapply(model_list, predict, newdata = test_dataB))


test_preds$log_PFS <- test_dataB$log_PFS


test_results <- test_preds |> 
  rename(MARS = y) |> 
  pivot_longer(cols = -log_PFS, names_to = "Model", values_to = "Predicted")  |> 
  group_by(Model)  |> 
  summarize(
    Test_RMSE = sqrt(mean((log_PFS - Predicted)^2)),
  )  |> 
  arrange(Test_RMSE) 

knitr::kable(test_results)
```
